---
title: "Market Model"
format: html
---

```{r}
#| warning: false
#| echo: false
source('functions.R')
```

To determine the normal return we can adopt the market model [@MacKinlay1997]
$$
D_{it} = \alpha_i + \beta_i D_{gt} + \epsilon_{it}
$$

And the AR is 
$$
AR_{it} = D_{it} – (\alpha_i + \beta_i D_{gt})
$$

Where:

$D_{it}$ is *`r credit_card.target_label()`* (differenced) $i$ at time $t$, \
$\alpha_i$ is the firm-specific intercept, \
$\beta_i$ is the sensitivity of firm value $i$ to the group $g$,\
$D_{gt}$ is the group *`r credit_card.target_label()`* (differenced) at time $t$, and
$\varepsilon_{it}$ is the error term for $i$ at time $t$.

## Model Summary

```{r}
#| warning: false
data <- get_model_data() |> drop_na(UBPRE524.diff)

estimation_data <- data |> filter(is.na(Partnership)) |> fill_gaps()
observation_data <- data |> filter(!is.na(Partnership))

results <- tibble(model = character(),
                    rsme = numeric())

market_formula <- "UBPRE524.diff ~ UBPRE524.group.diff"
m_table <- estimation_data |> fabletools::model(tslm = TSLM(as.formula(market_formula)))
m_table |> tidy()
```

```{r}
#| warning: false
#| label: fig-plot-fit
#| fig-cap: "Fitted vs. Observed for estimation data on sub-set of firms"

augmented_data <- m_table |> augment()

augmented_data |> filter(BankName %in% unique(observation_data$BankName)) |>
        autoplot(UBPRE524.diff, colour = "darkgrey") + 
        geom_line(aes(y = .fitted, colour = "#D55E00")) + facet_wrap(~BankName+BankType, ncol = 2) +
        labs(title = "<span style='color:#D55E00'>Fitted</span> vs. <span style='color:darkgrey'>Observed</span>") +
        theme(legend.position = "none", plot.title = element_markdown(),plot.subtitle = element_markdown())

```


### Visual Checks

```{r}
#| warning: false
#| label: fig-fit
#| fig-cap: "Linear regression model diagnostics"
#| layout-ncol: 3
#| fig-subcap:
#|   - "Residuals vs Fitted - The red line is not flat suggesting we need to add additional interaction terms to better capture the relationship between the predictors and the response variable"
#|   - "Q-Q Residuals - the S shape indicates residuals have heavier tails than expected normal distribution"
#|   - "Scale-Location - slight funnel shape indicates heteroscedasticity i.e. violates linear assumption of equal variance"
augmented_data|> 
  ggplot(aes(x = UBPRE524.diff, y = .fitted)) +
  geom_point() +
        labs(y = "Fitted (predicted values)",
             x = "Data (actual values)",) +
        theme(legend.position = "none", plot.title = element_markdown(),plot.subtitle = element_markdown())+
  geom_abline(intercept = 0, slope = 1, colour="red")
  
qqnorm(augmented_data$.fitted, pch = 1, frame = FALSE);qqline(augmented_data$.fitted, col = "steelblue", lwd = 2)

augmented_data|> group_by(BankName) |> mutate(y = sqrt(abs(.fitted/sd(.fitted)))) |>
  ggplot(aes(x = UBPRE524.diff, y = y)) +
  geom_point() +
        labs(y = "sqrt standardised residuals",
             x = "Data (actual values)",) +
        theme(legend.position = "none", plot.title = element_markdown(),plot.subtitle = element_markdown())
        

```

### Cross-Validation

```{r}
#| warning: false
#m_table |> filter(grepl("BARCLAYS",BankName)) |> select(tslm) |> gg_tsresiduals()

#A portmanteau test returns a large p-value, also suggesting that the residuals are white noise.
augment(m_table) |>
  filter(grepl("BARCLAYS",BankName))  |>
  features(.innov, ljung_box)  

bar_est_data <- estimation_data |> filter(grepl("BARCLAYS",BankName))
bar_event_data <- observation_data |> filter(grepl("BARCLAYS",BankName))
bar_data <- estimation_data |> filter(!is.na(UBPRE524.diff)) |> add_row(observation_data)
m_table |>
    filter(grepl("BARCLAYS",BankName))  |>
    forecast(new_data=bar_event_data) |>
    filter(.model=='tslm') |> View()
    autoplot(bar_data, colour = "pink")  + 
    geom_line(aes(x = Quarter, y=UBPRE524.diff), data= bar_event_data)



# selected_model <- m_table |> filter(grepl("BARCLAYS",BankName))
# fabletools::glance(selected_model) |> arrange(selected_model) |> select(.model:BIC)

# bar_est_data |> 
#        run_timeseries_cv(selected_model$tslm)

  
```
